---
title: "Ames H2o"
output: html_notebook
---
In this notebook I will be going through some model building using the Ames housing dataset in R.  In this scenario, I want to perform two different tasks, supervised regression and dimensionality reduction.  

### Package Import:

```{r, messages = FALSE}
library(rsample)  # data splitting 
library(dplyr)    # basic data manipulation procedures
library(ggplot2)  # plotting
library(h2o)
library(tibble)
# h20 setup
h2o.init()
```

### Data splitting:

```{r}
# Create training (70%) and test (30%) sets for the AmesHousing::make_ames() data.
# Use set.seed for reproducibility

set.seed(123)
ames_split <- initial_split(AmesHousing::make_ames(), prop = .7, strata = "Sale_Price")
ames_train <- training(ames_split)
ames_test  <- testing(ames_split)
```

### Glmnet model:

```{r}
# convert data to h2o object
ames_h2o <- ames_train %>%
  mutate(Sale_Price_log = log(Sale_Price)) %>%
  as.h2o()

# set the response column to Sale_Price_log
response <- "Sale_Price_log"

# set the predictor names
predictors <- setdiff(colnames(ames_train), "Sale_Price")


# try using the `alpha` parameter:
# train your model, where you specify alpha
ames_glm <- h2o.glm(
  x = predictors, 
  y = response, 
  training_frame = ames_h2o,
  nfolds = 10,
  keep_cross_validation_predictions = TRUE,
  alpha = .25
  )

# print the mse for the validation data
print(h2o.mse(ames_glm, xval = TRUE))

# grid over `alpha`
# select the values for `alpha` to grid over
hyper_params <- list(
  alpha = seq(0, 1, by = .1),
  lambda = seq(0.0001, 10, length.out = 10)
  )

# this example uses cartesian grid search because the search space is small
# and we want to see the performance of all models. For a larger search space use
# random grid search instead: {'strategy': "RandomDiscrete"}

# build grid search with previously selected hyperparameters
grid2 <- h2o.grid(
  x = predictors, 
  y = response, 
  training_frame = ames_h2o, 
  nfolds = 10,
  keep_cross_validation_predictions = TRUE,
  algorithm = "glm",
  grid_id = "ames_grid", 
  hyper_params = hyper_params,
  search_criteria = list(strategy = "RandomDiscrete")
  )

# Sort the grid models by mse
sorted_grid <- h2o.getGrid("ames_grid", sort_by = "mae", decreasing = FALSE)
# grab top model id
best_h2o_model <- sorted_grid@model_ids[[1]]
best_model <- h2o.getModel(best_h2o_model)
```


```{r}
coefs = data.frame(h2o.coef_norm(best_model)) %>% rownames_to_column() %>% arrange(desc(h2o.coef_norm.best_model.))

coefs %>% head()
```

Random Forest:

The basic algorithm for a regression random forest can be generalized to the following:

1.  Given training data set
2.  Select number of trees to build (ntrees)
3.  for i = 1 to ntrees do
4.  |  Generate a bootstrap sample of the original data
5.  |  Grow a regression tree to the bootstrapped data
6.  |  for each split do
7.  |  | Select m variables at random from all p variables
8.  |  | Pick the best variable/split-point among the m
9.  |  | Split the node into two child nodes
10. |  end
11. | Use typical tree model stopping criteria to determine when a tree is complete (but do not prune)
12. end

```{r}
# hyperparameter grid
hyper_grid.h2o <- list(
  ntrees      = seq(200, 500, by = 150),
  mtries      = seq(15, 35, by = 10),
  max_depth   = seq(20, 40, by = 5),
  min_rows    = seq(1, 5, by = 2),
  nbins       = seq(10, 30, by = 5),
  sample_rate = c(.55, .632, .75)
)

# random grid search criteria
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "mse",
  stopping_tolerance = 0.005,
  stopping_rounds = 10,
  max_runtime_secs = 30*60
  )

# build grid search 
random_grid <- h2o.grid(
  algorithm = "randomForest",
  grid_id = "rf_grid2",
  x = predictors, 
  y = response, 
  training_frame = train.h2o,
  hyper_params = hyper_grid.h2o,
  search_criteria = search_criteria
  )

# collect the results and sort by our model performance metric of choice
grid_perf2 <- h2o.getGrid(
  grid_id = "rf_grid2", 
  sort_by = "mse", 
  decreasing = FALSE
  )
print(grid_perf2)
```


